Foregrounds in Wide--Field Redshifted 21~cm Power Spectra

### 2015-02-23 ms.tex
Remove commented text
from 049772a:
============

In ms.tex,

(*) Remove commented text

%% \title{Detecting Epoch of Reionization in redshifted 21~cm: contamination in EoR window}
%% \title{Limits on the detection of the Epoch of Reionization from MWA observations of the redshifted 21~cm line}
\title{Foregrounds in Wide--Field Redshifted 21~cm Power Spectra}

@@

At the end of the recombination epoch, the Universe was completely neutral. This period, referred to as the {\it Dark Ages} in the Universe's history, is characterized by the localized accumulation of matter under the influence of gravity. And it ended with the formation of the first stars and galaxies which started emitting ultra--violet and X--ray radiation, thereby reionizing the neutral medium in their surroundings. This commenced the epoch of reionization (EoR) --- a period of non--linear growth of matter density perturbations and astrophysical evolution. % Studying the EoR holds the key to understanding this evolution.

@@

% {\bf We investigate signatures generic to all wide--field measurements of EoR power spectra and explore antenna shapes that are most desirable from the viewpoint of foreground contamination for future EoR experiments}. After matching observations from the Murchison Widefield Array with instrumental simulations of the entire sky, we examine the {\bf shape of the delay spectrum} of each baseline. We report two important findings: in the {\it avoidance} strategy, the foregrounds that most severely {\bf obscure} the redshifted 21~cm power spectrum are not caused by emission in the central field of view, but rather by bright objects from near the horizon; and, diffuse Galactic emission plays a significant role hitherto unpredicted. We quantify these by separating the simulations into components based on type of foreground emission. We then arrive at a new method for minimizing the contribution of bright foregrounds which uses prior knowledge of the sky to down--weight adversely contaminated baselines.

@@

% The presence of a ground plane below the dipole and the phased array introduces a smoother suppression of sensitivity near the horizon, whereas this transition is abrupt in the case of the dish.

@@

% \subsubsection{Diffuse Foreground Model}\label{sec:DSM}

@@

% \subsubsection{Compact Foreground Model}\label{sec:CSM}

@@

% \subsection{The Murchison Widefield Array (MWA)}\label{sec:instrument}

MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. The MWA is a 128--tile interferometer capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. See \citet{bea12} for the tile layout. %{\bf At the observing frequency of 185~MHz, we use $A_\textrm{e}=10.5$~m$^2$}.

The MWA passband of width $\Delta B=30.72$~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter. The coarse channel shape is obtained using an 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$. Each of these coarse bands consists of 32 fine channels of width 40~kHz each. After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level. % Any spectral flags, or other known weights constitutes $W_\textrm{i}(f)$ in equation~\ref{eqn:obsvis}.

% \subsection{EoR Observations}\label{sec:obsparms}

@@

% {\bf We estimate the system temperature, $T_\textrm{sys}$, using the {\it rms} of $\tilde{V}_b(\tau)$ obtained after {\it delay--deconvolution} across all antenna spacings for $|\tau| \geq 1$~$\mu$s using the relation:
% \begin{align}\label{eqn:Tsys}
%   \Delta \tilde{V} &= \frac{2\,k_\textrm{B}\,T_\textrm{sys}\sqrt{\Delta B}}{A_\textrm{e}\sqrt{2\,\Delta t}} 
% \end{align}
% where, $\Delta f = 80$~kHz, and $\Delta t = 112$~s. The choice of threshold for $\tau$ is well outside the {\it foreground window}, where foreground contamination is negligible and thus yields a robust estimate of $T_\textrm{sys}$. We find the average system temperature to be $\sim 95$~K across all frequency channels and across all antenna pairs throughout the course of the observation. Hence, for our simulations, we use $T_\textrm{sys}=95$~K to match the thermal noise observed in the data.}

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the delay power spectra plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively. The horizon delay limits are shown as a pair of vertical dotted lines. % The gray shaded area denotes the envelope of expected uncertainty in the {\bf delay power spectrum}. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the % $\gtrsim$~70\% error owing to 
% uncertainty in predicting the spectral index of compact foreground model, while thermal noise fluctuations dominate outside.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figure9.eps}
\includegraphics[width=\linewidth]{figure8.eps}
\caption{Simulated delay power spectra for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}). The baseline length and orientation is specified in each panel. The solid red, cyan, and black lines denote contributions from diffuse, compact, and composite foreground models respectively. Vertical dotted lines mark the horizon delay limits. Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a characteristic {\it pitchfork}--shaped structure. % The shaded region denotes the envelope of uncertainty around the composite model delay power spectrum. This uncertainty is dictated by uncertainties in spectral indices of foreground models inside the delay horizon delay limits and by thermal noise fluctuations outside.
\label{fig:pitchfork-baselines}}
\end{figure}

@@

% Note that since {\it visibility} in equation~\ref{eqn:obsvis} is a complex quantity, its {\bf delay} transform {\bf contains} positive and negative delays. Negative and positive delays correspond to the two hemispheres of the sky as transected by the antenna spacing vector. 
Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper. $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site. Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at $\tau>0$ and $\tau<0$, respectively. Similarly, an object in the northern sky will appear at $\tau>0$ for an antenna spacing oriented northward. For all observations used in this study, we use zenith as the phase center, for which $\tau\equiv 0$.

>>>>

\title{Foregrounds in Wide--Field Redshifted 21~cm Power Spectra}

@@

At the end of the recombination epoch, the Universe was completely neutral. This period, referred to as the {\it Dark Ages} in the Universe's history, is characterized by the localized accumulation of matter under the influence of gravity. And it ended with the formation of the first stars and galaxies which started emitting ultra--violet and X--ray radiation, thereby reionizing the neutral medium in their surroundings. This commenced the epoch of reionization (EoR) --- a period of non--linear growth of matter density perturbations and astrophysical evolution.

@@

@@

@@

@@

@@

MWA construction was completed in 2012 and, after commissioning, began its EoR observing program in 2013. The MWA is a 128--tile interferometer capable of observing a 30.72~MHz instantaneous band anywhere in the range 80--300~MHz. Each tile is a phased array of 16 dipoles, each in the shape of a bow--tie. This yields a primary field of view $\gtrsim$~20\arcdeg~ wide and multiple secondary lobes. See \citet{bea12} for the tile layout.

The MWA passband of width $\Delta B=30.72$~MHz is divided coarsely into 24$\times$1.28~MHz sub--bands with each sub--band weighted by a digital filter. The coarse channel shape is obtained using an 8--tap polyphase filter bank (PFB) and a Kaiser window with parameter $\beta=5$. Each of these coarse bands consists of 32 fine channels of width 40~kHz each. After correcting for the shape of these coarse channels, the fine channels at the edges of these sub--bands are flagged because they are known to be contaminated by aliasing at a low level.

@@

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the delay power spectra plots shown in Figure~\ref{fig:noiseless-dsm-csm-delay-spectrum}. The diffuse, compact, and composite components are shown as solid red, cyan, and black lines, respectively. The horizon delay limits are shown as a pair of vertical dotted lines.

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{figure9.eps}
\includegraphics[width=\linewidth]{figure8.eps}
\caption{Simulated delay power spectra for three chosen northward oriented antenna spacings of length: $\sim$84~m ({\it top}), $\sim$104~m ({\it middle}), and $\sim$171~m ({\it bottom}). The baseline length and orientation is specified in each panel. The solid red, cyan, and black lines denote contributions from diffuse, compact, and composite foreground models respectively. Vertical dotted lines mark the horizon delay limits. Compact emission dominates the central regions of the delay power spectra while both components, especially diffuse emission on short antenna spacings, dominate near the horizon delay limits, giving rise to a characteristic {\it pitchfork}--shaped structure. \label{fig:pitchfork-baselines}}
\end{figure}

@@

Figure~\ref{fig:delay-cartoon} illustrates the radio interferometer delays and conventions used in our paper. $\boldsymbol{b}$ is assumed to be on a coordinate system aligned with the local east, north (along local meridian) and upward (zenith) directions at the telescope site. Hence, a perfectly eastward oriented antenna spacing will observe objects in the eastern and the western skies at $\tau>0$ and $\tau<0$, respectively. Similarly, an object in the northern sky will appear at $\tau>0$ for an antenna spacing oriented northward. For all observations used in this study, we use zenith as the phase center, for which $\tau\equiv 0$.
